---
title: "Head Calibration and VOR Gain"
author: "Adam Pallus"
date: "2 April 2018"
output:
  html_document:
    df_print: paged
---

In this report, we examine a data file that shows signs of containing miscalibrated data.

The file is gN73AS3.txt

###Examine calibration phase during first 12000 data points:
```{r,echo=FALSE,message=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE,cache=FALSE)
```

```{r,message=FALSE,echo=FALSE,warning=FALSE}
library(signal)
library(plotly)
library(ggplot2)
library(dplyr)
source('knighthelperfunctions.R')
samplerate<- 304.7508/1000

```


```{r,eval=FALSE}

path<- "F:/dropbox/kdata/testanti/"
h<-loadnewheadfree(path=path)

ha<- AdjustCalibration(h,headoffset=3)
```

```{r firstplot}
ggplotly(ggplot(filter(ha,time<13000))+
           geom_line(aes(time,HV/10),color='blue',linetype=2)+
           geom_line(aes(time,H),color='blue')+
           geom_hline(yintercept = c(-60,60))+
           ylab('Head Position (deg)'))
```


```{r skip50}

ggplotly(ggplot(filter(h,time>50,time<13000))+
           geom_line(aes(time,H,label='Original Head Position'),color='blue')+
           geom_line(aes(time,E),color='darkred')+
           geom_line(aes(time,cumsum(HV*1)/samplerate/1000,label='Recalculated Head Position'),color='orange')+
           geom_hline(yintercept = c(-60,60))+
           ylab('Head Position (deg)'))
```


```{r}
ggplotly(
  h %>%
    filter(time>50,time<13000) %>%
    mutate(Hvcalc=cumsum(HV)/samplerate/1000,
           Gvcalc=Hvcalc+E) %>%
    ggplot()+
    geom_line(aes(time,Gvcalc),color='darkgreen')+
    geom_line(aes(time,Hvcalc),color='blue')+
    geom_hline(yintercept = c(-60,60))+
    ylab('Position (deg)')+
    ggtitle('Positions calculated from Velocity'))
```


This is almost perfect but there is still a little offset to the left. Maybe we can fix this by adding a small offset and reducing the gain a bit?

```{r}
ggplotly(
  h %>%
    filter(time>500,time<13000) %>%
    mutate(Hvcalc=cumsum(HV*1)/samplerate/1000+3,
           Gvcalc=Hvcalc+E) %>%
    ggplot()+
    geom_line(aes(time,Gvcalc),color='darkgreen')+
    geom_line(aes(time,Hvcalc),color='blue')+
    geom_hline(yintercept = c(-60,60))+
    ylab('Position (deg)')+
    ggtitle('Head Gain 1 and offset 3 deg right'))
```

```{r scrap}
htest<-h



htest %>%
  dplyr::filter(time>500) %>%
  dplyr::mutate(Hcalc=cumsum(HV*1)/samplerate/1000+3,
         Gcalc=Hcalc+E)->
  htest

ggplotly(
  htest %>%
    filter(time>nrow(htest)-13000)%>%
    ggplot()+
    geom_line(aes(time,Gcalc),color='darkgreen')+
    geom_line(aes(time,Hcalc),color='blue')+
    geom_hline(yintercept = c(-58,58),color='red')+
    geom_line(aes(time,Targ),size=2,alpha=0.5)+
    ylab('Position (deg)')+
    ggtitle('Positions calculated from Velocity'))

```

```{r}

filterButter<- function(y,freqs=0.0005,type='high'){
  require(signal)
  
  bf <- butter(2, freqs,type=type)
  # bf <- butter(2, c(.0005,.006))
  return(filtfilt(bf, y))
}
```

```{r}
ggplotly(
  htest %>%
    mutate(Hfilt=filterButter(Hcalc,freqs=0.00007),
           Gfilt=Hfilt+E)%>%
    filter(time>nrow(htest)-13000)%>%
    ggplot()+
    geom_line(aes(time,Gfilt),color='darkgreen')+
    geom_line(aes(time,Hfilt),color='blue')+
    geom_hline(yintercept = c(-58,58),color='red'))

ggplotly(
  htest %>%
    mutate(Hfilt=filterButter(Hcalc,freqs=0.00007),
           Gfilt=Hfilt+E)%>%
    filter(time<13000)%>%
    ggplot()+
    geom_line(aes(time,Gfilt),color='darkgreen')+
    geom_line(aes(time,Hfilt),color='blue')+
    geom_hline(yintercept = c(-58,58),color='red'))
```




ggplot(htest)+
  # geom_line(aes(time,Gcalc),color='darkgreen')+
  geom_point(aes(time,Hcalc),color='black',alpha=0.5)+
  geom_hline(yintercept = 0,color='red')+
  ylim(-10,10)

htest %>%
  mutate(Hfilt=filterButter(Hcalc,freqs=0.0005))->
  htest

ggplot(htest)+
  # geom_line(aes(time,Gcalc),color='darkgreen')+
  geom_point(aes(time,Hfilt),color='black',alpha=0.5)+
  geom_hline(yintercept = 0,color='red')+
  ylim(-10,10)

ggplot(dplyr::filter(htest,time<12000))+
  # geom_line(aes(time,Gcalc),color='darkgreen')+
  geom_line(aes(time,Hcalc),color='blue',alpha=0.5)+
  geom_line(aes(time,Hfilt+E),color='darkgreen',alpha=0.5)+
  geom_hline(yintercept = c(-58,0,58),color='red')
```

